---
title: "How Distributions Work"
output: html_document
---
Stan uses distributions to approximate answers using probabilities instead of using math rules. A very simple statement like "42 = variableA" trivially should assign to variable A the value 42. We are going to do it a different way with sampling. 

Introducing the normal distribution, a.k.a. your friend/nemisis the bell curve.

```{stan output.var=the_answer_normal}
parameters {
  real<lower=0,upper=3> sigma_standard_dev;
  real<lower=0,upper=100> the_answer;
}

model {
  42 ~ normal(the_answer,sigma_standard_dev);
}

```

```{r}
library(rstan);

fit <-
  sampling(the_answer,iter=1000, chains=1)

```
The normal distribution has the nice property that the mean, 'the_answer', is interpretable as the one data point we provided it. Below is graph of the normal distribution.

<show histogram>

```{stan output.var=the_answer_exponential}

parameters {
  real<lower=0,upper=10> lambda_the_answer_exponential;
}

model {
  42 ~ exponential(lambda_the_answer_exponential);
}
```

```{r}
library(rstan);

fit <-
  sampling(the_answer_exponential,iter=1000, chains=4)

```
The mean and 95% interval of the_answer_exponential bears no resemblelance to 'the_answer', but the the_answer_exponential does configure, or paramaterize the exponential distribution that we then can simulate data from to see where '42' lay. 
