---
title: "How Distributions Work"
output: html_document
---
Stan uses distributions to approximate answers using probabilities instead of using math rules you are familliar with from basic algebra. A very simple statement like "42 = variableA" trivially should assign to variable A the value 42. We are going to do it a different way with sampling. 

Introducing the normal distribution, a.k.a. your friend/nemisis the bell curve.

```{stan output.var=the_answer_normal}
parameters {
  real<lower=0,upper=3> sigma_standard_dev;
  real<lower=0,upper=100> the_answer;
}

model {
  42 ~ normal(the_answer,sigma_standard_dev);
}

```

```{r}
library(rstan);

fit <-
  sampling(the_answer,iter=1000, chains=1)

hist(extract(fit)$the_answer)


```
The normal distribution has the nice property that the mean, 'the_answer', is interpretable as the one data point we provided it. Below is graph of the normal distribution.

<show histogram>

```{stan output.var=the_answer_lognormal}

parameters {
  real<lower=0,upper=5> log42;
  real<lower=0,upper=10> sigma_standard_deviation;
}

model {
  42 ~ lognormal(log42,sigma_standard_deviation);
}


```

```{r}
library(rstan);

fit <-
  sampling(the_answer_lognormal,iter=1000, chains=4)

hist(extract(fit)$log42)
print(fit)
exp(mean(extract(fit)$log42))
```
The mean and 95% interval of the_answer_exponential bears no resemblelance to 'the_answer', but the the_answer_exponential does configure, or paramaterize the exponential distribution that we then can simulate data from to see where '42' lay. 
```{stan output.var=the_answer_lognormalx20}

parameters {
  real<lower=0,upper=5> log42x20;
  real<lower=0,upper=10> sigma_standard_deviationx20;
}

model {
  for(i in 1:20){
    42 ~ lognormal(log42x20,sigma_standard_deviationx20);
  }
}
```

```{r}
library(rstan);

fit <-
  sampling(the_answer_lognormalx20,iter=1000, chains=4)

print(fit)
exp(mean(extract(fit)$log42))

```

Data always wins if there is enough of it. Perhaps only for point estimates. Check with Jonathan.

```{stan output.var=}
data {
  int <lower=0> N;
  real <lower=-100,upper=100> uniform_generated_data[N];
}

parameters {
  real<lower=-100,upper=100> mu_mean;
  real<lower=0,upper=100> sigma_std_deviation;
}

model {
  uniform_generated_data ~ uniform(mu_mean,sigma_std_deviation);
}

  
```



