---
title: "Stan Is The Plan"
author: "Jonathan Auerbach, Breck Baldwin"
date: "8/17/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Course Outline

- Programmer focused on mechanics, not statistical theory
- Enough statistics to get you started understanding what is going on overall
- Get past high probability failure points

```{r fig.height=3}
hist(rexp(100,1),xlab="Stan Knowledge in Days",ylab="% Chance of Failure",main="Liklihood of Failure to Learn Stan")
```

## The State of Science

* Averages are old school
  + Easy science has been done, broad strokes of reality sorted
  + Population pooled theory worked out with N's in the 1,000s
* Human understandability of modeling 
  + Deep Learning utterly opaque
  + Do you really want black box AI doing prison sentencing?
* Most of science is small to medium data
  + Clinical trials in the 10's to 100's of subjects
  + Big Data approaches fail

## The Bayesian Modeling Response

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

